---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-", 
  warning = FALSE, 
  message = FALSE
)
```

# lexiquer

This package is a wrapper around Lexique 3.81, a database for natural language processing in French. 

More info on: [http://www.lexique.org](http://www.lexique.org)

## What's in this db?

Lexique gives access to ~ 150 000 french words with several annotations: lemme, phoneme, genre, frequency, number of letters, word neighbours... 

## Getting started

The full corpus is a data object contained inside the package, which you can call with :

```{r}
library(lexiquer)
data("lexique")
```

You can then left join it with a one-word-per-row data.frame: 

```{r message=FALSE}
library(tidytext)
library(proustr)
library(tidyverse)

sw <- proust_stopwords()
ad <- albertinedisparue

tm <- unnest_tokens(ad, word, text) %>%
  count(word) %>%
  anti_join(sw) 
tm %>%
  left_join(lexique, by = c("word" = "ortho")) %>%
  select(lemme, cgramortho) %>%
  na.omit() %>%
  count(lemme, cgramortho) %>%
  top_n(10, n) %>%
  arrange(desc(n))
```

### `bind_*` wrappers

`{lexiquer}` provides a series of wrapper to bind specific part of the corpus to your text. See the `bind_*` functions for more details. 

For example, you can binf the grammatical category of the word: 

```{r}
bind_gram_cat(tm, word)
```

### `is_lemme` 

Test if a word is a lemme : 

```{r}
is_lemme(tm, word)
```

### `count_*` wrappers

Several counting functions are available:

```{r}
count_syll(tm, word)
```


## Install 

```{r eval = FALSE}
devtools::install_github("ColinFay/lexiquer")
```

## Feedbacks 

Questions and feedbacks [welcome](mailto:contact@colinfay.me)!

You want to contribute ? Open a [PR](https://github.com/ColinFay/lexiquer/pulls) :) If you encounter a bug or want to suggest an enhancement, please [open an issue](https://github.com/ColinFay/lexiquer/issues).


